---
output: 
  github_document:
    toc: TRUE
title: "ggkids"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev = "ragg_png", # <- This piece specifically
  out.width = "100%"
)
```

<!-- badges: start -->
<!-- badges: end -->

Work builds on [2025-01-28-ggkids](https://evamaerey.github.io/mytidytuesday/2025-01-28-ggkids/ggkids.html) and [daphne](https://github.com/EvaMaeRey/daphne) package (now depricated).

The goal of ggkids is to ...

## Installation

You can install the development version of ggkids from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("EvaMaeRey/ggkids")
```

## Example


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "ragg_png", # <- This piece specifically
  dpi = 400, # <- This piece specifically
  fig.width = 12, 
  fig.height = 6,
  warning = F,
  message = F
)
```




ggkids is about bringing the grammar of graphics and ggplot2 to very young learners.

ggplot2 is quite logical and accessible, but may present some challenges for young learners given a more limited vocabulary and little experience interpreting error messages. 


0. the package contains data that let young learners replicate several plots from the new 'Daphne Draws Data' book by Storytelling with Data Founder Cole Nussbaumer-Knaflic and illustrator John Skewes, a couple plots from Josh Starmer's book 'Norm and 'Squatch Make a New Friend'.


1. `ggkids()` is an alias for `ggplot()`.  It can be used to initialized the plot and set global data just like `ggplot()`, but `ggkids` might be preferred if you want to more strongly signal 'this version is just for you, kiddos!'. But brings along it's own theme. 

```{r}
library(ggplot2)
```

<details>


```{r theme_kids}
#' @export
theme_kids <- function(...){
  ggplot2::theme_classic(
    paper = "whitesmoke", 
    ink = "darkgrey", 
    base_size = 30,
    base_family = "Comic Sans MS") 
}

# todo... move to theme
update_geom_defaults(GeomPoint, aes(size = from_theme(pointsize * 3)))

update_geom_defaults(GeomText, aes(size = from_theme(pointsize * 3)))

```


```{r}
ggplot(cars) + 
  aes(speed, dist) +
  geom_point() + 
  geom_text(label = "🎈") +
  theme_kids() 
```

```{r, ggkids}
ggkids <- function(data, 
                   spec_theme = theme_kids()){
  
 ggplot2::ggplot(data = data) + spec_theme
       
      }
```

```{r}
ggkids(cars) + 
  aes(speed, dist) + 
  geom_point()
```

</details>

```{r, echo = F}
"ggkids(mtcars) + 
  aes(cyl) + 
  geom_bar() + 
  labs(title = 'Hello ggkids!')" |>
  ggram::ggram(code = _, "ggkids brings it's own theme...")
```






2. `use` is an alias for `aes`, since 'aesthetic' is unlikely to be in the vocabulary of young learners.  We think that you might say 'let's use color to communicate about height' (use the color aesthetic channel to communicate about 'var1') or 'use size' might be a natural way to talk about aesthetic mapping with children.


3. A series of `use_*()` functions are also available and facilitate more step-wise plot construction, e.g. `use_x()`, `use_y()`, `use_size()`, `use_linewidth()`  🚧 *One thing people struggle with including me is the color aesthetic.  It sometimes feels like fill might better be called color, and color something (outline_color?); If this is done using open point (shape = 21), would probably be the way to go too.*

<details>


```{r aes_to_use}
#' @export
encode <- function(color, ...){
  aes(color = {{color}}, fill = {{color}}, ...) 
}

#' @export
use <- function(color, ...){
  aes(color = {{color}}, fill = {{color}}, ...) 
}

#' @export
use_x <- function(x){list(aes(x = {{x}}))}

#' @export
use_y <- function(y){list(aes(y = {{y}}))}


#' @export
use_weight <- function(weight){aes(weight = {{weight}})}

#' @export
use_area <- function(area){aes(weight = {{area}})}

#' @export
use_rows <- function(rows, cols, ...){facet_grid(rows = vars({{rows}}), cols = vars({{cols}}), ...)}

#' @export
use_columns <- function(cols, rows, ...){facet_grid(rows = vars({{rows}}), cols = vars({{cols}}), ...)}

#' @export
use_rows_columns <- function(rows, cols, ...){facet_grid(rows = vars({{rows}}), cols = vars({{cols}}), ...)}

#' @export
use_wrap <- function(wrap, ...){facet_wrap(facets = vars({{wrap}}), ...)}

#' @export
use_size <- function(size){aes(size = {{size}})}

#' @export
use_shape <- function(shape){aes(shape = {{shape}})}

#' @export
use_color <- function(color){aes(fill = {{color}})}

#' @export
use_color_line <- function(color){aes(color = {{color}})}

#' @export
use_chart_point <- function(...){qlayer(geom = qproto_update(GeomPoint, aes(shape = 21), 
                                                         required_aes = c()),
                                    stat = qstat(function(data, scales){data$x <- data$x %||% 0 ; data$y <- data$y %||% 0; data}), ...)}


# data <- function(data){ggplot(data |> remove_missing()) + theme_classic(ink = "darkgrey", paper = "whitesmoke", base_size = 18)}

# chart_jitter <- geom_jitter

# chart_heat <- function(...){list(
#   qlayer(geom = GeomTile, 
#          stat = qproto_update(StatSum, aes(fill = after_stat(n), size = NULL)), ...),
#   scale_fill_gradientn(colors = c("blue", "white", "yellow", "orange", "red")),
#   theme(panel.grid.minor = element_line(color = "darkgrey")))
# }


# 
# title <- function(title){labs(title = title)}
# subtitle <- function(subtitle){labs(subtitle = subtitle)}
# caption <- function(caption){labs(caption = caption)}
# tag <- function(tag){labs(tag = tag)}
```

</details>


4. More aesthetic defaults are provided, rather than required for layer rendering.  e.g. to add a point layer, `x` and `y` are not required, but instead default to zero. We think inspecting a plot with a zero default might be more useful than reading an error message for very young learners.  🚧 *This is done using the helper function, `aes_default`. *
5. New `picture` aesthetic will play more of a role - because kids love pictures, and faces, etc.   `stamp_picture()` is lets you add an annotation layer in an arbitrary place on your plot, e.g. `stamp_picture(🚀, x = 4, y = 10)` - position default is the middle of the plot, i.e. `x = I(.5), y = I(.5)` and picture default is a classic smiley  🙂.
5. We introduce `chart_*()` series which can be used similar to layer functions (`geom_*` and `stat_*`), however they bring defaults that are appropriate for the chart type in question, e.g. `chart_plunging_bar()` wraps up `geom_bar(stat = "identity")` with `scale_y_reverse()`.  Available chart functions: `chart_bar`, `chart_item_stack`
6. Some of the chart functions are supported with new Stats, which are StatItemStack, StatPointCount, 



<details>


```{r}
library(statexpress)
library(tidyverse)

# update_geom_defaults(GeomPoint, aes(size = from_theme(pointsize * 3)))

#' @export
aes_default <- function(default = aes(x = 0)) {

  structure(
    list(
         default_spec = default), 
    class = "aes_default"
    )

}


#' @import ggplot2
#' @importFrom ggplot2 ggplot_add
#' @export
ggplot_add.aes_default <- function(object, plot, object_name) {
  
  if(is.null(plot$mapping[[names(object$default_spec)]])){
   plot <-  plot + object$default_spec
  }

  plot

}

```




```{r}
#' @export
stamp_picture <- function(picture = "🙂", x = I(.5), y = I(.5), ...){
  
  annotate(geom = GeomText, label = picture, x = x, y = y, ...)
  
}

```

</details>

## Pets Popularity

<details>

```{r chart_pie}
#' @export
chart_pie <- function(...){
  
  list(
  
  geom_bar(position = "fill", width = 1, show.legend = F, ...),
  # add defaults that bar doesn't usually include
  aes_default(aes(y = .5)), 
  aes_default(aes(fill = "All")),
  aes_default(aes(weight = 1)),
  aes_default(aes(color = from_theme(paper))),
  # add labels
  stat_count(geom = GeomLabel, color = "transparent",
             position = position_fill(vjust = .5),
             aes(label = after_stat(fill), 
                 y = .9, 
                 # group = after_stat(fill)
                 ), alpha = 0,
             show.legend = F,
             size = 30
             ),
  coord_polar(),
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank()),
  labs(fill = NULL)
  
  )
  
}
```




```{r}
pets_data <- data.frame(pets = c("🐱", "🐶", "🦚", "🐠", "🐰"), 
                   number_of_pets = c(30, 25, 10, 15, 5)) |> 
  mutate(pets = fct_infreq(pets, number_of_pets) |> fct_rev())


usethis::use_data(pets_data, overwrite = T)

```



```{r}
pets_data

pets_data |> 
  ggplot() +
  chart_pie() +
  use_color(pets) + 
  use_area(number_of_pets)
```


</details>

```{r, echo = F}
'pets_data <- 
  tribble(~pets, ~number_of_pets,
          "🐱",  30, 
          "🐶",  25, 
          "🦚",  10, 
          "🐠",  15, 
          "🐰",   5)
        
# Then she drew the data ...        
ggkids(data = pets_data) +
  chart_pie() +
  use_color(pets) + 
  use_area(number_of_pets) ' |>
ggram::ggram("Daphne🐲📊 tabulates the data and draws the data!", 
      code = _, widths = c(1,1.6))





```



# Crustaceans Race

<details>


```{r crustaceans}
library(tidyverse)
types <- c("🦐", "🦀")
set.seed(1234)
ocean_table <- cars |> 
  rename(size = dist) |>
  mutate(type = c(
    rep("🦐", 20),
    sample(types, 10, replace = T),
    rep("🦀", 20))) |>
  sample_frac()
```


```{r GeomPointFill}
#' @export
GeomPointFill <- qproto_update(GeomPoint, aes(shape = 21),
                              required_aes = c())


library(statexpress)
chart_point <- function(...){
  list(stat_identity(geom = GeomPointFill, show.legend = F, ...),
  aes_default(aes(x = 0)),
  aes_default(aes(y = 0)),
  # aes_default(aes(shape = I(after_stat(picture)))),
  scale_size(range = c(2,10))
  
  )
  }


chart_picture <- chart_point

# should replace with lm xy
chart_fit_global_line <- function(...){
  
  geom_smooth(method = lm, ..., show.legend = F, se = F, 
              linetype = "dashed",
              aes(shape = NULL, 
                  picture = NULL,))
  
}

use_picture <- function(picture){aes(shape = I({{picture}}))}
```




```{r}
head(ocean_table)

ocean_table |>
  ggkids() + 
  use_y(speed) +
  use_x(size) + 
  chart_point() +
  use_picture(type) + 
  use_size(size) + 
  chart_fit_global_line()
```

</details>


```{r, echo = F}
"ggkids(ocean_table) + 
  use_y(speed) +
  use_x(size) + 
  chart_point() +
  use_picture(type) + 
  use_size(size) + 
  chart_fit_global_line() + 
  labs(x = 'little         big') +
  labs(y = 'slow           fast')" |>
ggram::ggram(code = _, "Are crabs 🦀🦀🦀 faster than shrimp 🦐🦐🦐?",
             subtitle = "Size explains speed", widths = c(1, 1.6))
```




# Fry Eating Contest

(Norm and 'Squatch Make a New Friend - Joshua Starmer)

<details>


```{r}
real_fries_table <- data.frame(id_fry = 1:7, 
                               seconds = c(1, 1, 2, 1, 2, 3, 1) |> as.character(),
                               height =  c(1, 2, 1, 3, 2, 1, 4),
                               item = rep("🟨",7))


all_fries_table <- data.frame(id_fry = 1:10, 
                              seconds = c(1, 1, 2, 1, 2, 3, 1, 1,2,3)  |> as.character(),
                              height =  c(1, 2, 1, 3, 2, 1, 4, 5, 3, 2),
                              item = c(rep("🟨", 7), 
                                        rep("🟩", 3),
                              ind_real = c())
                              )

imaginary_fries_table <- all_fries_table |> slice(8:10)
```


```{r}
chart_tile <- function(...){
  
  list(
    qlayer(geom = qproto_update(GeomTile, aes( width = .75, height = .75)), ...),
    coord_equal()
    
  )
  
}


set_color <- function(color){
  
  aes(fill = I(color |> alpha(.5)) , color = I(color))
  
}

```


```{r}
real_fries_table |>
  ggkids() + 
  use(x = seconds, 
      y = height, 
      picture = item) + 
  chart_point() 

last_plot() %+% all_fries_table

real_fries_table |>
  ggkids() + 
  use(x = seconds, y = height) + 
  chart_tile() + 
  set_color("goldenrod2")

last_plot() +
  chart_tile(
    data = imaginary_fries_table,
    set_color("darkseagreen4")
    )
```


</details>


```{r, echo = F}
'real_fries_table |>
  ggkids() + 
  use_x(seconds) + 
  use_y(height) + 
  chart_tile() + 
  set_color("goldenrod2") +
  chart_tile(
    data = imaginary_fries_table,
    set_color("darkseagreen4")
    )' |> 
  ggram::ggram(code = _, title = "Norm and Squatch and the Fry Eathing Contest")

```




# Bananas Gone Missing

<details>


```{r jungle_table}
theme_chart_bar <- function(){
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_blank())
  }

chart_bar <- function(...){
  list(geom_col(...), 
       theme_chart_bar(),
       scale_y_continuous(expand = expansion(c(0, .3))),
       labs(x = NULL))
}

compute_item_stack <- function(data, scales, width = 0.2){
               
    data$shape <- data$shape %||% data$picture

    data |> 
      uncount(y) |>
      dplyr::mutate(row = row_number()) |> 
      dplyr::mutate(y = row - 
        0.5) |>
      dplyr::mutate(width = width)
    
  }

chart_item_stack <- function(...){
  
  list(
  qlayer(
    geom = GeomPointFill, 
    stat = qstat(compute_item_stack)
  ),
  # spacing
  qlayer(
    geom = GeomTile, 
    stat = qstat(compute_item_stack), 
    alpha = 0
  ),
  scale_y_continuous(expand = expansion(c(0, .3))),
  aes_default(aes(y = 1)),
  aes_default(aes(x = "All")),
  labs(x = NULL, y = NULL),
  guides(y = "none")
  )

  
}

ggprop.test:::compute_group_bricks

jungle_table <- data.frame(tree = paste0("🌴#", 1:5), 
                     num_bunches = c(2, 5, 1, 2, 1), 
                     banana = "🍌")

jungle_table |> 
  select(x = tree, y = num_bunches, picture = banana) |>
  compute_item_stack()
```

</details>


```{r, include = F}
'
jungle_table <- 
  tribble(~tree, ~num_bunches, 
          "🌴#1", 2,
          "🌴#2", 5,
          "🌴#3", 1,
          "🌴#4", 2,
          "🌴#5", 1)

jungle_table |> 
  ggkids() + 
  use_x(tree) + 
  use_y(num_bunches) + 
  use_picture("🍌") + 
  chart_item_stack()' |>
ggram::ggram(code = _, widths = c(1,1.6))
```

```{r, echo = F}
last_plot() + 
  stamp_picture("🐒",
                size = 40,
                x = I(.95),
                y = I(.15)) + 
  coord_cartesian(clip = "off")
```

# See, Touch, Hear, Smell, Taste

```{r, echo = F}
'
senses_table <- 
  tribble(~order, ~sense, ~reps, 
          "1st", "👁️", 5,
          "2nd", "✋", 4,
          "3rd", "👂", 3,
          "4th", "👃", 2,
          "5th", "👅", 1)

senses_table |> 
  ggkids() + 
  use_x(order) + 
  use_y(reps) + 
  use_picture(sense) + 
  chart_item_stack()' |>
ggram::ggram(code = _, widths = c(1,1.6))



```

<details>


```{r}

real_fries_table <- data.frame(id_fry = 1:7, 
                               seconds = c(1, 2, 1, 1, 2, 3, 1),
                               height =  c(1, 1, 2, 3, 2, 1, 4))


all_fries_table <- data.frame(id_fry = 1:10, 
                               seconds = c(1, 2, 1, 1, 2, 3, 1, 1,2,3),
                               height =  c(1, 1, 2, 3, 2, 1, 4, 5, 4, 2))
# last_plot() + 
#   annotate(geom = GeomText,
#            x = I(.75), y = I(.72),
#            label = "🎈🎀🙏",
#            angle = -10,
#            size = 22,
#             )  
```


</details>



```{r}
'ggkids(jungle_table) + 
  encode(x = tree,
      y = num_bunches) + 
  chart_bar() +
  set_color("lemonchiffon2")' |>
  ggram::ggram(code = _, 
               title = "chart_bar in ggkids")


```




# Polar Bears Dive


<details>


```{r}
chart_bar_plunging <- function(...){
  list(geom_label(vjust = 1, aes(label = after_stat(y), fill = NULL),
                  linewidth = 0),
    geom_col(...), 
       theme_chart_bar(),
       scale_y_reverse(expand = expansion(c(.3, 0))),
       scale_x_discrete(position = "top"),
       labs(x = NULL)
       )
}
```


```{r}
bears <- paste0("🐻‍❄️ #", 1:5)
depth <- c(3,5,4,6,5)

polar_bear_table <- data_frame(bears, depth)

usethis::use_data(polar_bear_table, overwrite = T)

```


</details>


```{r, echo = F}
'tribble(~id_bear, ~depth,
            "🐻‍❄#1",    3,
            "🐻‍❄#2",    5,
            "🐻‍❄#3",    4,
            "🐻‍❄#4",    6,
            "🐻‍❄#5",    5) |>
ggkids() +
  use_x(bears) + 
  use_y(depth) +
  chart_bar_plunging() + 
  set_color("lightblue")' |>
  ggram::ggram(code = _,
               title = "Daphne went to the north pole...")

```

# Digging Instruments


<details>


```{r}
time <- c(0, 15, 30, 45, 60)
num_tunnels <- c(0, 4, 5, 5.5, 5.9)
fork_and_spoon_table <- data_frame(time, num_tunnels, type = "🍴")

shovel_and_bucket_table <- data_frame(time, num_tunnels = num_tunnels * 2, type = "🪣")

paws_table <- data_frame(time, num_tunnels = num_tunnels * 3, type = "🐾")

chart_line <- geom_line

digging_table <- fork_and_spoon_table |>
  bind_rows(shovel_and_bucket_table) |>
  bind_rows(paws_table)
```




</details>

```{r, echo = F}
'ggkids(digging_table) +
  use_x(time) + 
  use_y(num_tunnels) + 
  chart_line() +
  # use_group() +
  chart_point() + 
  use_picture(type) +
  labs(x = "⏱️ (minutes)",
       y = "Number of Tunnels")' |>
  ggram::ggram(code = _, title = "What's the best for digging")
```



# Outer Space Travel

(future work)

<details>

```{r shuttles}

outer_space_data <- data.frame(shuttle = paste0("🚀#", 1:6), fuel = c(.3,.5,.3, .8,.7, .4))

chart_portion_full <- function(...){
  
  list(
       geom_col(fill = "transparent", aes(y = 1)),
       geom_col( ... ),
       aes_default(aes(color = from_theme(ink)))
       )
  
}

stamp_hline <- function(y = .5, linetype = "dashed", ...){
  
  geom_hline(yintercept = y, linetype = linetype, ...)
  
}
```

</details>

```{r, echo = F}
outer_space_data

'outer_space_data |>
ggkids() +
  use_x(shuttle) +
  use_y(fuel) +
  chart_portion_full() +
  set_color("darkolivegreen3") +
  stamp_hline(.75) + 
  stamp_picture("👽",
                angle = 25,
                size = 40,
                x = I(.95),
                y = I(.15))' |>
  ggram::ggram(code = _ , title = "Which aliens can get home?")
```

# Scooter Repair Cost

<details>

```{r}
weeks <- c(0,1,2)
num_coins <- c(1,2,3)
coin <- c("🪙", "🪙🪙", "🪙🪙🪙")

scooter_table <- tibble(weeks, num_coins, coin)

usethis::use_data(scooter_table, overwrite = T)
```


```{r}
compute_panel_count <- function(data, scales){
            data |> 
             uncount(y, .remove = F)
           }

scooter_table |> 
  select(x = weeks, y = num_coins) |> 
  compute_panel_count()

chart_point_count <- function(...){
  list(
  qlayer(geom = GeomPointFill,
         stat = qstat(compute_panel = compute_panel_count), 
         position = position_jitter(width = .12, height = .12),
         ..., show.legend = F),
  aes_default(aes(x = 0)),
  aes_default(aes(y = 0)),
  aes_default(aes(shape = I(after_stat(picture)))),
  scale_size(range = c(2,10))
  
  )
  }
```

</details>

```{r, echo = F}
set.seed(12345)

'repair_data <- 
   tribble(~weeks, ~num_coin,
            0,      1,
            1,      2,
            2,      3)
            
ggkids(data = repair_data) +
  use_x(weeks) + 
  use_y(num_coins) + 
  chart_point_count() + 
  use_picture("🪙") +
  chart_line() +
  stamp_hline(2.25) + 
  stamp_picture("🛴",
                y = 2.45,
                x = 1.9,
                size = 25,
                angle = 10)' |> 
  ggram::ggram(code = _, title = "Declan saves to repair his scooter")

```



